---
title: "cs624-hw2-ryan-richardson"
author: "Ryan Richardson"
date: "13/11/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(foreign)
library(dbplyr)
library(ROCR)
library(Metrics)
library(ResourceSelection)
```

## Pre-Processing

Start with basic preprocessing and drop the following:

Age is dropped due to being duplicated in agec
Bmi is dropped due to being calculated by height and weight
dibpat is dropped due to being duplicated by behpat
Id is droppepd due to being an identifier field not related to data
sbp is droppped due to being contained in lnsbp
typchd69 is dropped due to being directly related to our target chd
weight and weightcat are dropped due to being contained in lnwght
smoking is dropped due to being contained in ncigs
Arcus is converted to a categorical variable
time169 is removed for being contained in our target variable
Records with NA values are dropped as there is a minimal number
```{r preprocess}
wcgs <- read.dta('wcgs.dta')
wcgs$arcus = factor(wcgs$arcus)
filtered = wcgs[, -c(1,4,7,8,10,14,15,17,18,20,21)]
filtered <- filtered[complete.cases(filtered),]
summary(filtered)
```

## Backward Elimination Main Effects Model Building

We start with a general linear model using chd69 as our variable of interest, and all other variables as predictors
Models are shown with the following variables dropped over time:
Behbat
t1
height
uni
arcus

The model we are left with shows a an almost% chance to have some form of CHD, for someone with a 
Cholesterol of 0, systolic bp of exp(0), a weight of exp(0), smoking 0 cigarretees a day, under
age 40.

For every 1 point in cholesterol a person is 1% more likely to experience CHD, for every 
std.dev they are above the mean sbp they 14% more likely to experience CHD. For every std.dev
they are above the mean in weight they are 5.5% more likely to experience CHD). For every 
cigarette they smoke per day they are 1% more likely to experience CHD. Someone in the age category 41-45 is 0.8%
more likely to experience CHD, someone 51-55 is 1.9% more likely to experience CHD, and someone from 56-60 is 2.6% more
likely to experience CHD than the baseline. 


Once we plot the ROC Curve, we get the best AUC of 0.959 when using a threshold of 0.88

Based on the threshold values needed to attain this AUC, I would be worried about the accuracy of the model. 
Primarily because such a high threshold is dangerously close to simply saying that every person will have coronary heart
disease. Even so, we have a high AUC value, which tells me that this model is missing something very significant, and that
some effects are missing from it. 

The Hosmer-Lemeshow GoF test gies an extremely small p-value, close to 0, indicating a poorly fit model
```{r mainEffectsModelbuilding, echo=FALSE}
mbest <- filtered[,-c(1, 2, 5, 9, 10)]
mbest = mbest[complete.cases(mbest),]
mainModelBest <- glm(chd69~., data=mbest, family="binomial")
summary(mainModelBest)
mainPredictions = predict(mainModelBest, type="response")
mainPredictionTest = prediction(mainPredictions, mbest$chd69)
mainperf <- performance(mainPredictionTest, measure = "tpr", x.measure ="fpr")
plot(mainperf)

mainBestAuc = 0
mainBestAucThreshold = 0
for (i in seq(min(mainPredictions),1,0.001)){
  mainPredThresh = ifelse(mainPredictions > i, 1, 0)
  mainCurrAuc = auc(mainPredThresh, mbest$chd69)
  if (!is.nan(mainCurrAuc) && mainCurrAuc >= mainBestAuc){    
    mainBestAuc = mainCurrAuc
    mainBestAucThreshold = i
  }
}

mainHosmerTest <- hoslem.test(mbest$chd69, fitted(mainModelBest), g = 100)
print(mainHosmerTest)

print("Main Effects Model Best AUC")
print(mainBestAuc)
print("Main Effects Model Best AUC Threshold")
print(mainBestAucThreshold)

```

## Backward Elimination Main & 2-Way Interaction Effects Model Building

Starting again with our main model using chd69 as our variable of interest
Lnsbp was left in as was weight due to being close to the 0.05 significance level
Agec was retained as well for having significant interactions
Models are shown with the following dropped over time:
uni
lnsbp
t1
behbat
height
ncigs
arcus

This model is looking at someone with a choelsterol of 0, of average weight, age less than 40 to begin with.
There are significant interactions between weight and cholesterol seen in this model. 
For most of the included effects, every point increase will result in a substantial increase 
in CHD risk rate. There is a surprising drop in CHD chance for peopl age 56-60 off the bat, but that 
seems to be partly due to the interaction between lnwght and that age group causing a substantial increase
in risk. Beacuse of this, it may have been a good idea to remove the age category from the model and 
replace it with  some other age measure. 


When we plot the ROC curve we get the best AUC of 0.96042 at a threshold of 0.5

I'm much happier with the threshold values giving us the AUC in this model. This indicates not only 
a decent prediction, but the lower threshold indicates a much lower chance that we classify every as 
having a CHD event or not.

Unfortunately, the Hosmer-Lemeshow goodness of fit still indicates a poorly fit model. This may 
be due to missing factors, high covariance in the data, or some other mispecification caused by the 
target variables used. Such as the agec. 
```{r mainInteractionEffectsModelbuilding, echo=FALSE}
best <-filtered[,-c(1, 2, 5, 6, 8, 9, 10)]
best = best[complete.cases(best),]
interactionModelBest <- glm(chd69~.^2, data=best, family=binomial(link = "logit"))
summary(interactionModelBest)
predictions = predict(interactionModelBest, type="response")
predictionTest = prediction(predictions, best$chd69)
perf <- performance(predictionTest, measure = "tpr", x.measure ="fpr")


bestAuc = 0
bestAucThreshold = 0
for (i in seq(min(predictions),1,0.001)){
  predThresh = ifelse(predictions > i, 1, 0)
  currAuc = auc(predThresh, best$chd69)
  if (!is.nan(currAuc) && currAuc >= bestAuc){    
    bestAuc = currAuc
    bestAucThreshold = i
  }
}
plot(perf)
HosmerTest <- hoslem.test(best$chd69, fitted(interactionModelBest), g = 10)
print(HosmerTest)

print("Interaction Effects Model Best AUC")
print(bestAuc)
print("Interaction Effects Model Best AUC Threshold")
print(bestAucThreshold)
```
