---
title: "Homework 3"
author: "Ryan Richardson"
date: "10/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MASS)
library(VGAM)
```

## Question 1


```{r preprocess}
summary(Boston)
houseData = Boston
houseData$chas = as.factor(houseData$chas)
medianThreshold = as.vector(quantile(houseData$medv, c(0.25, 0.75)))
houseData$medianValue = cut(houseData$medv, c(0, medianThreshold[1], medianThreshold[2], max(houseData$medv)), c("low", "medium", "high"))
levels(houseData$medianValue)
```


```{r buildModel, warning=FALSE}

baseModel <- vglm(medianValue~., family=multinomial, data = houseData[,-c(14)])
step4vglm(baseModel)

bestModel <- vglm(formula = medianValue ~ crim + indus + nox + rm + age + 
    dis + rad + tax + ptratio + black + lstat, family = multinomial, 
    data = houseData[, -c(14)])

bestSummary <- summary(bestModel)
bestSummary
```

```{r odds}

odds <- exp(bestSummary@coef3[,1])
odds
```

The base model was for High value housing.


The default intercepts for other categories are close to 0.00, signalling a very low 
odds ratio of being in the High value class by default. 

For each additional point in crime, you were 6% more likely to be in Low price housing over high, but you were 15% less likely to be in Medium Price Housing than High Price.

For each additional point increase in Indus you were 1.8% more likely to be in Low Price housing and 13.4% more likely to be in Medium Price Housing than High Price.

For each additional point in NoX Emissios you were 1.6 x 10^8 times more likely to be in Low Cost Housing, and 6.63 x 10^2 times more likely to be in Medium Price Housing than High Price.

For each additional room you were 84% less likely to be in Low Price and 87% less likely to be in Medium Price Housing than High Price. 

For each additional year  older your house was, you were 3.8% more likely to be in Low Price Housing and 0.4% less likely to be in Medium Price Housing than High Price.

For each additional unit of distance from an employment center you were 164% more likely to be in Low price and 65% more likely to be in Medium Price Housing than High Price. 

For each additional unit in radial distance from a highway you were 37% less likely to be in low price and 25% less likely to be in Medium Price Housing than High Price. 

For each additional point in property tax rates you were 1.6% more likely to be in Low Price and 1.0% more likely to be in Medium Price Housing than High Price. 

For each additional unit in pupil-teach ration you were 120% more likely to be in Low Price and 49.5% more likely yo be in Medium Price Housing than High Price. 

For each additional point in proportion of black population you were 0.02% less likely to be in low price and 0.4% more likely to be in Medium Price Housing than High Price. 

For each % lower in status of the population you were 82% more likely to be in low price and 43% more likely to be in Medium Price Housing than High Price. 
